\section{Requirements from Language Specifications}\label{sec:spec}

The XMP Fortran language specification~\cite{xmp} supports many of 
the coarray features defined in the Fortran~2008 standard~\cite{coarray}, 
and intrinsic procedures {\tt CO\_SUM}, {\tt CO\_MAX}, {\tt CO\_MIN}, and 
{\tt CO\_BROADCAST} defined in the Fortran~2018 standard~\cite{coarray18}are supported.
In addition, the XMP C language specification was extended to support coarray features.

This section introduces the coarray features and what is required
of the compiler in order to implement the coarray features.


%-----------------------------------------------------------------------------
\subsection{Images Mapped to XMP Nodes}\label{sec:spec-image}
%-----------------------------------------------------------------------------

In the Fortran standard, an {\bf image} is defined as an instance of a program. 
Each image executes the same program and has its own individual data.
Each image has a different image index $k$.
While the Fortran standard itself does not specify where each image is executed, 
XMP specifies that images are mapped to executing nodes on a one-to-one basis.
Therefore, image $k$ is always executed on executing node $k$, where $1 \leq k \leq n$, and 
$n$ is the number of images as well as the number of executing nodes. 
Since each MPI rank number of {\tt MPI\_COMM\_WORLD} (0-origin) is 
always mapped to an XMP node number in order, image $k$ corresponds to 
rank $(k - 1)$.

Note that the executing nodes can be a subset of the entire (initial) node set. 
For example, two distinct node sets can execute two coarray subprograms concurrently.
The first executing images at the start of the program are entire images.
Coarray features are compatible with those of the Fortran standard, unless 
the {\tt TASK} and {\tt END TASK} directives are used.
If the execution encounters a {\tt TASK} directive specified with a subset of nodes, 
then the corresponding subset of images will be the executing images for the task region. 
The current number of images and my image number,  which are given by inquire functions
{\tt num\_images} and {\tt this\_image} also match the executing images, and
the {\tt SYNC\_IMAGES} statement synchronizes the executing images.
When the execution encounters the {\tt END TASK} directive corresponding to the
{\tt TASK} directive, the set of executing images is reinstated.

%   Coarray features can be used inside the TASK directive blocks. As default,
%   each coarray image is mapped one-to-one to a node of the current executing 
%   task. I.e., num_images() returns the number of nodes of the current executing 
%   task and this_image() returns each image index in the task.
%      There are two directives to change the default rule above. A COARRAY 
%   directive corresponding to a coarray declaration changes the image index set 
%   of the specified coarray with the one of the specified nodes. An IMAGE 
%   directive corresponding to one of a SYNC ALL statement, a SYNC IMAGES 
%   statement, a call statement calling CO_SUM, CO_MAX, CO_MIN or CO_BROADCAST 
%   changes the current image index set with the one of the specified nodes.
%   See the language spacifications [3].

\requirement
The runtime library should manage the executing image set and the current image index 
in a stack in order to reinstate them at the exit point of the task.


%-----------------------------------------------------------------------------
\subsection{Allocation of Coarrays}\label{sec:spec-coarray}
%-----------------------------------------------------------------------------

A {\bf coarray} or a coarray variable is a variable that can be referred from the other images. 
A coarray with the {\tt ALLOCATABLE} attribute is called an {\bf allocatable coarray}, and is
otherwise called a non-allocatable coarray. A non-allocatable coarray may not be a pointer 
and must have an explicit shape and the {\tt SAVE} attribute. In order to help 
intuitive understanding, we refer to a non-allocatable coarray a {\bf static coarray}. 
The lifetime of a static coarray is throughout execution of the program on all images, even if
the coarray is declared in a procedure called with a subset of images.

On the other hand, an allocatable coarray is allocated with the {\tt ALLOCATE} statement and 
freed either explicitly with the {\tt DEALLOCATE} statement or implicitly at the end of the 
scope in which the {\tt ALLOCATE} statement is executed ({\bf automatic deallocation}).

Static coarrays can be declared as scalar or array variables as follows:
\begin{verbatim}
      real(8), save :: a(100,100)[*]
      type(user_defined_type), save :: s[2,2,*]
\end{verbatim}

The square brackets in the declaration distinguish coarray variables from 
other (non-coarray) variables. The declaration declares the virtual shape of the images, and the last 
dimension must be deferred (as `{\tt *}').

Allocatable coarrays can be declared as follows:
\begin{verbatim}
      real(8), allocatable :: b(:,:)[:]
      type(user_defined_type), allocatable :: t[:,:,:]
\end{verbatim}


A notable constraint is that at any synchronization point in program execution, 
coarrays must have the same dimensions (sizes of all axes) for all images
({\bf symmetric memory allocation}). 
Therefore, a static coarray must have the same shape for all images during 
program execution, and an allocatable coarray must be allocated and deallocated 
collectively at the same time with the same dimensions for the executing images.
Thanks to the symmetric memory allocation rule, all executing images can have
the same symmetrical memory layout, which makes it possible to calculate the address 
of the remote coarray with no prior inter-image communication.

\requirement
Static coarrays must be allocated and made accessible remotely
before execution of the user program and 
made inaccessible remotely and be freed after execution of the user program.
In contrast, allocatable coarrays must be allocated and made accessible remotely
when the {\tt ALLOCATE} statement is encountered and 
made inaccessible remotely and freed when the {\tt DEALLOCATE} statement is encountered or 
the exit point of the scope to which the corresponding {\tt ALLOCATE} statement is encountered.


%-----------------------------------------------------------------------------
\subsection{Communication}\label{sec:spec-comm}
%-----------------------------------------------------------------------------

Coarray features in XMP include three types of communications between images, i.e.,
reference and definition to remote coarrays,
collective communications (intrinsic subroutines {\tt CO\_SUM}, {\tt CO\_MAX}, 
{\tt CO\_MIN}, and {\tt CO\_BROADCAST}), and
atomic operations ({\tt ATOMIC\_DEFINE} and {\tt ATOMIC\_REF}).
%
Collective communications and atomic operations are similar to those
in MPI library.
Communications for reference and definition to remote coarrays are 
characteristic for coarray features.


%- PUT communication
PUT communication is caused by an assignment statement with a {\bf coindexed variable} 
as the left-hand side expression, e.g.,
\begin{verbatim}
      a(i,j)[k] = alpha * b(i,j) + c(i,j)
\end{verbatim}
This statement causes the PUT communication to the array element {\tt a(i,j)}
on image {\tt k} with the value of the left-hand side.
%
Using the Fortran array assignment statement, array-to-array PUT communication 
can be written easily, e.g., the following statement causes 
{\tt M}$\times${\tt N}-element PUT communication:
\begin{verbatim}
      a(1:M,1:N)[k] = alpha * b(1:M,1:N) + c(1:M,1:N)
\end{verbatim}

%- GET communication
GET communication is caused by referencing the {\bf coindexed object}, 
which is represented by a coarray variable with cosubscripts enclosed by square brackets, 
e.g., {\tt s[1,2]} and {\tt a(i,j)[k]}, where {\tt s} and {\tt a} are scalar and 
two-dimensional array coarrays, respectively.
%
A coindexed object can appear in almost any expression, including array expressions.

\requirement
In order to implement definition/reference to a coindexed variable/object,
PUT/GET one-sided communication is suitable for use.
%
In order to avoid costly processing, such as a remote procedure call, 
remote direct memory access-based (RDMA-based) implementation is desirable.
%
In PUT/GET communication for large data, redundant multiple memory copies 
should carefully be avoided for all software layers, the communication library, the runtime, the Fortran library, and the object.

% A reference of coindexed object is basically converted to a runtime library call
% to get the result of GET communication. 
% Note that the result can be a large array.
% For example, in in array assignment statement:
% \begin{verbatim}
%       c(1:M,1:N) = a(1:M,1:N)[k] + b(1:M,1:N)
% \end{verbatim}
% coindexed object $a(1:M,1:N)[k]$ should be converted to a function that
% returns an array value shaped $[M, N]$.


%-----------------------------------------------------------------------------
\subsection{Synchronization}\label{sec:spec-sync}
%-----------------------------------------------------------------------------

%-- between images
The access order of coarrays between images is explicitly controlled by the 
programmer using the {\bf image control statement}, 
such as {\tt SYNC ALL} and {\tt SYNC IMAGES} statements. 
The statement allows the compiler system to make PUT/GET communication asynchronous.
The sequence of execution between the image control statements is called as a {\bf segment}.
An asynchronous communication must be completed by the end of the segment.

%-- inside image
Inside each image, the compiler must maintain data dependency 
as before, even if the compiler contains coarray communications.
The compiler suppresses the {\bf non-blocking communication},
which postpones waiting for communication completion.
In order to keep data dependency among the definitions and references to the same 
coarray in the same segment, non-blocking communication should be restricted.
The example below in which the same remote coarray is accessed a number of times inside the same segment.
\begin{quote}
\begin{verbatim}
 1      if (this_image()==1) then
 2          a[2]=
 3          =a[2]
 4          a[2]=
 5          a[2]=
 6      endif
\end{verbatim}
\end{quote}
Between lines 2 and 3, the completion wait for PUT communication is necessary in order
to avoid referencing data that is not defined completely.
Similarly, between lines 3 and 4, the completion wait for GET communication is 
necessary in order to avoid referencing data that is being updated.
However, between lines 4 and 5, the completion wait is not necessary.
The issue of race condition on image 2 cannot be avoided by the completion wait
on image 1 in general, and avoiding this issue is up to the programmer.

\requirement
%To reduce the latency overhead, non-blocking one-sided communication is effective.
%The compiler should generate non-blocking GET communication as long as possible.
Unless the same remote data is accessed from the same segment, 
non-blocking completion can be delayed until the end of the segment.
%However, the possibility to meet the condition above should 
%be took account. 
Since the data received by the GET communication is usually referenced soon, 
non-blocking GET communication is hard to use. Therefore, if GET communication is
always on blocking, then only the flow dependency (between lines 2 and 3) should be considered.


%-----------------------------------------------------------------------------
\subsection{Subarrays and Data Contiguity}\label{sec:spec-contig}
%-----------------------------------------------------------------------------

Except for a dummy argument, an array is fully {\bf contiguous} across the dimensions.
A subarray of the array can be fully or partially contiguous or non-contiguous.
For example, if an array is declared with the shape {\tt a(1:M,1:N)},
then the whole array (referenced as {\tt a} or {\tt a(:,:)} or {a(1:M,1:N)})
is fully contiguous and a subarray {\tt a(2:5,3)} is partially contiguous.
We defined a term {\bf contiguous length} as the length for which the data is partially
contiguous. For example, the contiguous lengths of {\tt a(2,3)} and {\tt a(2:5,3)} are
1 and 4, respectively. {\tt a(1:M,1:3)} is two-dimensionally contiguous and has 
contiguous length $2 \times {\tt M}$.
{\tt a(1:M-1,1:3)} is one-dimensionally contiguous and has a 
contiguous length $({\tt M} - 1)$.

\requirement
For high-performance communication, it is important to find the contiguous length
across the dimensions, because thousands of bytes of contiguous data is needed in order to be 
comparable to the communication latency in general, and only the first dimension 
of the array is not always long enough.
%
% a(i,j)[k]は代わりに全体配列にも部分配列にもなれるので，
% coindexed variableについても次元を跨いだ連続性の抽出が必要である．
% それに加えて、ローカル側、すなわち、右辺式データの連続性も意識に入れなければならない。
% 高速な通信を実現するには，左辺と右辺で共通に連続な区間を検出してその単位で通信を反復するか，
% 右辺データは連続区間にpackして左辺の連続区間を単位として通信を反復するなどの戦略がある。


%-----------------------------------------------------------------------------
\subsection{Coarray C Language Specifications}\label{sec:spec-c}
%-----------------------------------------------------------------------------

The XMP language specification extends the C language to support coarray 
features. Array notations, such as subarray and array assignment statements,
are adopted in the C language.
%
In XMP/C, a coarray is a data object but is not a pointer.
A coarray is either 1) of basic type, 2) a structure in which
no component is a pointer, or 3) an array of 1, 2, or 3.

XMP/C also has static and allocatable coarrays.
Coarray variables declared directly in the file and declared with 
the {\tt static} attribute are static.
Coarray variables can be allocated with intrinsic functions.

% Cのcoarrayは、通常のCの変数と同じように、引数渡しやcast演算によって自由に
% その型と形状の解釈を変えることができる。
% これらの仕様と制限は、Cプログラマ
% にとっての使いやすさを考えて、Cらしいプログラミングスタイルを認めた。
% Coarray C++とは違うアプローチである。

% cf.\ air:/Users/iwashita/Desktop/coarray/Project\_Coarray/の下にいくつか





